---
title: "Untitled"
author: "Kiera Matiska"
date: "10/23/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, options(round = 3))

# Add packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

# Read in urchins.csv and clean up column names

- Use the function read_csv() to read in the data
- Use the fuction clean_names() to change column headers to lower-snake case

```{r}
urchins <- read_csv(here("data", "urchins.csv"))

urchins_clean <- urchins %>% 
  clean_names()
```

# Section 1: Size Comparisons

### Data Wrangling

Use select() to select the columns size and common_name create a new data frame

```{r}
urchins_size <- urchins_clean %>% 
  select(common_name, size)
```


### Data visualization

Using ggplot(), create a histogram using geom_histogram to compare the distribution of sizes between red and purple urchins
- Reduce number of bins to ~10
- Add a caption figure

```{r}
ggplot(data = urchins_size,
       aes(x = size, color = common_name)) +
  geom_histogram() + # Try to figure out how to add color
  facet_grid(~ common_name) +
  stat_bin(bins = 10) +
  labs(title = "Size Distributions of Purple and Red Urchins",
      x = "Urchin Diameter (cm)",
      y = "Population Count")
```

Figure 1: Size Distributions of Purple and Red Urchins. Sizes are measured using centimeter on sea urchins within this sample.

### Statistical Analyisis

Using summarize() create a summary table for purple and red sea urchins. Table should include:
- mean
- median
- standard deviation
- sample size

Use kableExtra to finalize table for knitted document and add a table caption above the table

```{r}
urchin_size_sum <- urchins_size %>% 
  group_by(common_name) %>% 
  summarize(mean(size),
            median(size),
            sd(size),
            sample_size = n())

urchin_size_sum %>% 
  kbl(caption = "Statistical data, including mean, median, standard deviation, and sample size, of purple and red sea urchin sizes") %>% 
  kable_styling()
```

